---
title: "Homework 2"
author: "Natasha Woon"
date: "2026-10-04"
output: github_document
---

Load Packages
```{r, message = F}
library(tidyverse)
library(readxl)
```

# Problem 1 

Importing CSV File

```{r, message = F}
subway_df = 
  read_csv(file = "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           na = c("", "NA", ".")) |> 
  janitor::clean_names() |>
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, entrance_type, vending, ada) |>
  mutate(
    entry = recode_values(
      entry, 
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```

This dataset contains data about NYC subway transit lines. The variables include line, station_name, station_latitude, station_longitude, route 1, route 2, route 3, route 4, route 5, route 6, route 7, route 8, route 9, route 10, route 11, entry, entrance_type, vending, and ada. These variables describe the station line, the station name, where the station is situated on a map, the 11 different routes, vending information, and information about the station entrances, and ADA compliance. 

There were originally other variables in the dataset. I only selected the variables required, and left out the other variables from the final data frame. Some of the variables that were cleaned out included information about division of the stations, staffing hours, and station exit only information. After extracting and cleaning the variables of interest, there are 1868 rows and 32 columns in this data frame. I also recoded the entry variable from a character (YES/NO) variable into a logical (TRUE/FALSE) variable. 

No, the original data import is not tidy. The `route1` through `route11` variables are all in separate columns instead of being within one column but separated by row. This means that these columns are values and not variable names. 

*How many distinct stations are there?*

```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```

There are 465 distinct stations. 

*How many stations are ADA compliant?* 

```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow() 
```

There are 84 stations that are ADA compliant. 

*What proportion of station entrances / exits without vending allow entrance?*

```{r}
subway_df |> 
  filter(vending == "NO") |>
  nrow()

subway_df |> 
  filter(entry == TRUE, vending == "NO") |>
  nrow()
  
```

There are 183 stations without vending. Out of these 183 stations, 69 stations allow entrance. 


Reformatting the data so that route number and route name are distinct variables. 

```{r}
subway_tidy_df = 
subway_df |>
  mutate(across(c(route8, route9, route10, route11), as.character)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

*How many distinct stations serve the A train?*

```{r}
subway_tidy_df |>
  filter(route_name == "A") |>
  distinct(station_name, line) |>
  nrow()
```

60 stations serve the A train. 

*Of the stations that serve the A train, how many are ADA compliant?*

```{r}
subway_tidy_df |>
  filter(route_name == "A", ada == TRUE) |> 
  distinct(line, station_name) |>
  nrow() 
```

Of the stations that serve the A train, 17 are ADA compliant. 

# Problem 2

Read, clean, and name the wheel variables for the Mr Trashwheel dataset. Also utilize the mutate function to convert character variables to numeric variables. 

```{r, warning = F}
mr_trashwheel_df =
  read_excel(path = "data/202610 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             range = "A2:N742",
             na = c("", "NA", ".")) |>
  janitor::clean_names() |>
  mutate(sports_balls = as.integer(sports_balls),
         year = as.numeric(year),
         wheel_name = "mr_trash_wheel")
```

Read, clean, and name the wheel variables for the Professor Trashwheel dataset. Also utilize the mutate function to convert character variables to numeric variables.

```{r, warning = F}
prof_trashwheel_df = 
  read_excel(path = "data/202610 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             range = "A2:M139",
             na = c("", "NA", ".")) |>
  janitor::clean_names() |>
  mutate(wheel_name = "prof_trash_wheel",
         plastic_bottles = as.numeric(plastic_bottles))
```

Read, clean, and name the wheel variables for the Gwynnda dataset.

```{r, warning = F}
gwyn_trashwheel_df = 
  read_excel(path = "data/202610 Trash Wheel Collection Data.xlsx", 
             sheet = "Gwynnda the Good Wheel of the W", 
             range = "A2:L394",
             na = c("", "NA", ".")) |>
  janitor::clean_names() |>
  mutate(wheel_name = "gwyn_trash_wheel")
```

Now, we combine the datasets. 

```{r}
trash_wheel_combined = bind_rows(mr_trashwheel_df, prof_trashwheel_df, gwyn_trashwheel_df)
```

Once the three trash wheel datasets were combined, there are now `r nrow(trash_wheel_combined)` observations and `r ncol(trash_wheel_combined)` columns. The variables include `dumpster`, `month`, `year`, `date`, `weight_tons`, `volume_cubic_yards`, `plastic_bottles`, `polystyrene`, `cigaretts_butts`, `glass_bottles`, `plastic_bags`, `wrappers`, `sports_balls`, `homes_powered` and `wheel_name`. The key variables are essentially the point in time in which the data is collected, the type of trash collected, and the name of the wheel. 

For the available data, the total weight of trash collected by Professor Trash Wheel is `r trash_wheel_combined |> filter(wheel_name == "prof_trash_wheel") |> pull(weight_tons) |> sum()` tons. 

The total number of cigarette butts collected by Gwynda in June of 2022 is `r trash_wheel_combined |> filter(wheel_name == "gwyn_trash_wheel", month == "June", year == 2022) |> pull(cigarette_butts) |> sum() |> format(scientific = FALSE)`. 

# Problem 3

Import, clean, and tidy the dataset of baseline characteristics for Problem 3. 

```{r}
MCI_baseline_df = 
  read_csv(
    file = "data/MCI_baseline.csv",
    skip = 1, 
    na = c("", "NA", ".")
    ) |> 
  janitor::clean_names() |>
  mutate(
    sex = recode(sex, `1` = "Male", `0` = "Female"),
    apoe4 = recode(apoe4, `1` = "APOE4 carrier", `0` = "APOE4 non-carrier")) |>
  filter(is.na(age_at_onset) | age_at_onset>current_age) 
```

I imported the CSV file into R Studio. I then skipped the first line of the CSV file since it was details about the codebook. Then, I standardized the NA variables. I then re coded the `sex` and `apoe4` variables from numeric into character variables based on the codebook provided in line 1 of the CSV file. To ensure that participants who do not meet the stated inclusion criteria (no MBI at baseline) are excluded, I utilized the filter function to remove participants who had MCI onset before baseline. I included the NA participants since they have yet to have a MCI onset. 

There are `r nrow(MCI_baseline_df)` subjects and `r ncol(MCI_baseline_df)` variables. The variables include `id`, `current_age`, `sex`, `education`, `apoe4`, and `age_at_onset`. There were `r nrow(MCI_baseline_df)` participants who were recruited, of which `r filter(MCI_baseline_df, !is.na(age_at_onset)) |> nrow()` developed MCI. 

The average baseline age is `r MCI_baseline_df |> pull(current_age) |> mean()` years. 

The proportion of women in the study who is an APOE4 carrier is `r nrow(filter(MCI_baseline_df, sex == "Female", apoe4 == "APOE4 carrier"))`/`r nrow(filter(MCI_baseline_df, sex == "Female"))` or `r nrow(filter(MCI_baseline_df, sex == "Female", apoe4 == "APOE4 carrier")) / nrow(filter(MCI_baseline_df, sex == "Female"))`. 








